---
title: 'Business Data Analysis Capstone Project'
author: "Leopoldine Mirtil"
date: "2023-08-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### DATA SOURCE

The data used in this capstone task is from the 'Application of Data Analysis in Business with R Programming' guided project, offered through Coursera.com and taught by instructor Nilosree Sengupta


### SCENARIO

You are provided with a dataset of a departmental store. It contains details of products from May 2020, a period marked by Covid-19. Your manager wants you to find out investing in which products will be more profitable. Your objective is to analyse the patterns and trends of the products, and gather insights for 
strategic decision making. (At this level, you don’t need to make reports/recommendations now. You will build reports and make recommendations in future, with the knowledge you gained from this project.)



### PART 1: IMPORT DATA

#### I. LOAD PACKAGES

```{r, message=FALSE}
require(ggplot2)
library(corrplot)
library(ggcorrplot)
library(tidyverse)
```


#### II. LOAD DATASET

```{r}
# set directory
setwd("C:/Users/Leopoldine/Desktop/Mine/Coding Projects & Portfolio/Guided Projects/R Guided Project/Application of Data Analysis in Business with R Programming/Department Store Project")

#load data
store <- read.csv("FINAL DEPARTMENTAL STORE.csv")

# view data
head(store)
```


### PART 2: DATA MANIPULATION

#### I. TRANSFORM DATA

```{r}
# arrange data in desc order of profit & use first 360 rows
store1 <- arrange(store, desc(PROFIT))%>%
  slice_head(n=360) 

# view data 
head(store1)
```



#### II. STATISTICAL INTERPRETATION OF DATA

```{r}
# find average, max and min of profit by product category
store1 %>%
group_by(PRODUCT_CATEGORY)%>%
  summarise(AVG=mean(PROFIT), MAX=max(PROFIT), MIN=min(PROFIT)) %>%
  print(n=38)
```



### PART 3: DATA VISUALIZATION

#### I. COLUMN PLOT FOR AVERAGE_NET_PROFIT & COMPANY

```{r, message=FALSE}
store1 %>% 
  group_by(COMPANY, PRODUCT_TYPE) %>% 
  summarise(AVERAGE_NET_PROFIT=mean(NET_PROFIT)) %>%
  ggplot(aes(x=COMPANY, y=AVERAGE_NET_PROFIT, fill=PRODUCT_TYPE))+
  geom_col() +
  scale_y_continuous(labels = scales::comma) 
```



#### II. SCATTER PLOT FOR SELLING_PRICE & QUANTITY_DEMANDED

```{r}
store1 %>%
  ggplot(aes(x=QUANTITY_DEMANDED, y=SELLING_PRICE, color= PRODUCT_CATEGORY))+
  geom_point() +
  theme(legend.title = element_text(angle = 90), legend.title.align = 0.5) +
  guides(colour = guide_legend(title.position = "left", ncol=5)) + 
  theme(legend.position = "bottom") 
```




#### III. COLUMN CHART OF PROD TYPE & AVERAGE NET PROFIT

```{r}
store1 %>% 
  group_by(PRODUCT_TYPE) %>% 
  summarise(AVERAGE_NET_PROFIT=mean(NET_PROFIT, na.rm = TRUE)) %>%
  arrange(desc(AVERAGE_NET_PROFIT)) %>%
  ggplot(aes(x=fct_reorder(PRODUCT_TYPE, +AVERAGE_NET_PROFIT), y=AVERAGE_NET_PROFIT))+
  geom_col(fill="cadetblue3") +
  xlab("PRODUCT_TYPE") +
  scale_y_continuous(labels = scales::comma) +
  coord_flip()
```



#### IV. HISTOGRAM FOR PROFIT PERCENT AND PRODUCT TYPE

```{r}
store1 %>%
  ggplot(aes(x=PROFIT_PERCENT, fill=PRODUCT_TYPE))+geom_histogram(binwidth = 20)
```



#### V. PIE CHART FOR PRODUCT CATEGORIES OF TOP 3 PRODUCT TYPES & TOTAL NET PROFIT

```{r}
store1 %>%
  filter(PRODUCT_TYPE=="hygiene" | PRODUCT_TYPE=="baby product" |PRODUCT_TYPE=="household") %>%
  group_by(PRODUCT_CATEGORY) %>%
  summarise(TOTAL_NET_PROFIT=sum(NET_PROFIT)) %>%
  arrange(desc(PRODUCT_CATEGORY)) %>%
  mutate(percentage=round(TOTAL_NET_PROFIT*100/sum(TOTAL_NET_PROFIT))) %>% 
  mutate(y_pos = cumsum(percentage)-0.5*percentage) %>%
  ggplot(aes(x="", percentage, fill=PRODUCT_CATEGORY))+
  geom_bar(width = 1, stat = "identity", color= "white", alpha=.5)+
  coord_polar("y", start=0)+
  geom_text(aes(y=y_pos, label=paste0(percentage, "%")), color="black")+
  scale_fill_manual(values=rainbow(10))+ theme_void()
``` 



### PART 4: CORRELATION 

#### I. CORRELATION MATRIX OF THE COMPLETE DATASET

```{r}
store<-dplyr::select_if(store, is.numeric)
r<-cor(store, use="complete.obs")
round(r,2)
```




#### II. HEAT MAP OF THE CORRELATION MATRIX  

```{r}
ggcorrplot(r)
```


